---
title: "Kaggle - Analysis with Titatic Data"
author: "Isaac Yauri"
date: "Monday, July 20, 2015"
output: pdf_document
---

# Titanic: Machine Learning from Disaster

Source: https://www.kaggle.com/c/titanic

Predict survival on the Titanic (using Excel, Python, R, and Random Forests)

See best practice code and explore visualizations of the Titanic dataset on Kaggle Scripts. Submit directly to the competition, no data download or local environment needed!

The sinking of the RMS Titanic is one of the most infamous shipwrecks in history.  On April 15, 1912, during her maiden voyage, the Titanic sank after colliding with an iceberg, killing 1502 out of 2224 passengers and crew. This sensational tragedy shocked the international community and led to better safety regulations for ships.

One of the reasons that the shipwreck led to such loss of life was that there were not enough lifeboats for the passengers and crew. Although there was some element of luck involved in surviving the sinking, some groups of people were more likely to survive than others, such as women, children, and the upper-class.

In this challenge, we ask you to complete the analysis of what sorts of people were likely to survive. In particular, we ask you to apply the tools of machine learning to predict which passengers survived the tragedy.

##VARIABLE DESCRIPTIONS:

survival        Survival
                (0 = No; 1 = Yes)
pclass          Passenger Class
                (1 = 1st; 2 = 2nd; 3 = 3rd)
name            Name
sex             Sex
age             Age
sibsp           Number of Siblings/Spouses Aboard
parch           Number of Parents/Children Aboard
ticket          Ticket Number
fare            Passenger Fare
cabin           Cabin
embarked        Port of Embarkation
                (C = Cherbourg; Q = Queenstown; S = Southampton)

##SPECIAL NOTES:

Pclass is a proxy for socio-economic status (SES)
 1st ~ Upper; 2nd ~ Middle; 3rd ~ Lower

Age is in Years; Fractional if Age less than One (1)
 If the Age is Estimated, it is in the form xx.5

With respect to the family relation variables (i.e. sibsp and parch)
some relations were ignored.  The following are the definitions used
for sibsp and parch.

Sibling:  Brother, Sister, Stepbrother, or Stepsister of Passenger Aboard Titanic
Spouse:   Husband or Wife of Passenger Aboard Titanic (Mistresses and Fiances Ignored)
Parent:   Mother or Father of Passenger Aboard Titanic
Child:    Son, Daughter, Stepson, or Stepdaughter of Passenger Aboard Titanic

Other family relatives excluded from this study include cousins,
nephews/nieces, aunts/uncles, and in-laws.  Some children travelled
only with a nanny, therefore parch=0 for them.  As well, some
travelled with very close friends or neighbors in a village, however,
the definitions do not support such relations.
<br/>  
<br/> 

##We start loading the data sets from Kaggle.

```{r}
setwd("c:/EMC/Cursos/GitHub/Practice-With-Titanic/")
train <- read.csv("train.csv")
test <- read.csv("test.csv")
```

##Loading Libraries

```{r}
library(randomForest)
```

##Review Data

```{r}
summary(train)
summary(test)
```

##Analizing data

1 Age: We have many missing values on train and test datasets.  
2 Fare: We have 1 missing value on test dataset.  
3 Embarked: We have 2 observations with "blank" value on train dataset.  

##Random Forest

To submit our Prediction, we can use Random Forest with all variables, but we will recive a message error because some of them are imcomplete. So, we need to use only Pclass, Sex, Sibsp and Parch. 

```{r}
TitanicRF <- randomForest(as.factor(Survived) ~ Pclass + Sex + SibSp + Parch, 
                              data = train, nodesize = 100, ntree = 2444, importance = TRUE)
PredictRF <- predict(TitanicRF, newdata = test)
PredTest <- predict(TitanicRF, newdata=test, type="response")
MySubmission <- data.frame(PassengerID = test$PassengerId, Survived = PredTest)
write.csv(MySubmission, "Submission1.csv", row.names=FALSE)
```

But you will obtain and score of 0.77512 with this submission.

checking importance of variables

```{r}
varImpPlot(TitanicRF)
```

Looks like Pclass and Sex are the most important variables.

To improve a better score, we need to solve those missing values

##Inputting missing Fare value for test dataset

We need to set a guess price (Fare) for this observation.
Investigate about this person to figure out better what was the Fare.

```{r}
test[which(is.na(test$Fare)), ]
```

Looking prices for people with the same info.

```{r}
MissingFare <- median(train[train$Pclass == 3 & train$Embarked == "S" & train$Age > 50 & train$Sex == "male", "Fare"], na.rm = TRUE)
MissingFare
```

Setting missing Fare with the average of people with the same characteristics.

```{r}
test[153, "Fare"] <- MissingFare
```

Creating a new Random Forest adding Fare variable.

```{r}
TitanicRF <- randomForest(as.factor(Survived) ~ Pclass + Sex + SibSp + Parch + Fare, 
                              data = train, nodesize = 50, ntree = 2444, importance = TRUE)
PredictRF <- predict(TitanicRF, newdata = test)
PredTest <- predict(TitanicRF, newdata=test, type="response")
MySubmission <- data.frame(PassengerID = test$PassengerId, Survived = PredTest)
write.csv(MySubmission, "Submission2.csv", row.names=FALSE)
```

We receive 0.78469 with this submission.

```{r}
varImpPlot(TitanicRF)
```

##Inputting Embarked value on train dataset

checking what is the observation with missing Embarked variable

```{r}
train[train$Embarked == "", ]
```

Looking similarities

```{r}
table(train[train$Survived == 1 & train$Pclass == 1, "Embarked"])
```

we can inputting a value "S" (Southampton) because is a mayority.

```{r}
train$Embarked[c(62,830)] = "S"
train$Embarked <- factor(train$Embarked)
```

train$Embarked <- factor(train$Embarked)

```{r}
TitanicRF <- randomForest(as.factor(Survived) ~ Pclass + Sex + SibSp + Parch + Fare + Embarked, 
                              data = train, nodesize = 50, ntree = 2444, importance = TRUE)
PredictRF <- predict(TitanicRF, newdata = test)
PredTest <- predict(TitanicRF, newdata=test, type="response")
MySubmission <- data.frame(PassengerID = test$PassengerId, Survived = PredTest)
write.csv(MySubmission, "Submission3.csv", row.names=FALSE)
```

No improvement with Embarked.

```{r}
varImpPlot(TitanicRF)
```

##Trying to solve missing Age info.

If we check some names, we can observe that they have their Title embedded. For example "master"

"Master is an English honorific for boys and young men."

Checking all "masters" in training data.

```{r}
train[grep("Master.", train$Name, fixed = TRUE), c("Name", "Age")]
```

checking their ages


```{r}
summary(train$Age[grep("Master.", train$Name, fixed = TRUE)])
```

So, is a good idea setting the Median to those missing values.
And we can use their Title (and other variables) to set those missing values.

To do this, we need to create a new varible and find this characteristic inside the name (Miss, Mrs., Ms., Mme., for example.)

##Creating variable Title

for this case, we need to merge both sets, but considering number of variables.

```{r}
all_set <- rbind(train[, -2], test)

all_set$Title <- NA
all_set$Alias <- 0

all_set$Title[grep("Master.", all_set$Name, fixed = TRUE)] <- "Master."
all_set$Title[grep("Miss", all_set$Name, fixed = TRUE)] <- "Miss"
all_set$Title[grep("Mr.", all_set$Name, fixed = TRUE)] <- "Mr."
all_set$Title[grep("Mrs.", all_set$Name, fixed = TRUE)] <- "Mrs."
all_set$Title[grep("Rev.", all_set$Name, fixed = TRUE)] <- "Rev."
all_set$Title[grep("Don.", all_set$Name, fixed = TRUE)] <- "Don."
all_set$Title[grep("Dr.", all_set$Name, fixed = TRUE)] <- "Dr."
all_set$Title[grep("Major.", all_set$Name, fixed = TRUE)] <- "Major."
all_set$Title[grep("Jonkheer", all_set$Name, fixed = TRUE)] <- "Jonkheer"
all_set$Title[grep("Col.", all_set$Name, fixed = TRUE)] <- "Col."
all_set$Title[grep("Mme.", all_set$Name, fixed = TRUE)] <- "Mme."
all_set$Title[grep("Ms.", all_set$Name, fixed = TRUE)] <- "Ms."
all_set$Title[grep("Lady.", all_set$Name, fixed = TRUE)] <- "Lady."
all_set$Title[grep("Sir.", all_set$Name, fixed = TRUE)] <- "Sir."
all_set$Title[grep("Mlle.", all_set$Name, fixed = TRUE)] <- "Mlle."
all_set$Title[grep("Capt.", all_set$Name, fixed = TRUE)] <- "Capt."
all_set$Title[grep("the Countess.", all_set$Name, fixed = TRUE)] <- "the Countess"
# Dona is spanish. We need change to Miss.
all_set$Title[grep("Dona.", all_set$Name, fixed = TRUE)] <- "Miss"
all_set$Alias[grep("(", all_set$Name, fixed = TRUE)] <- 1

# Set a factor for this new variable
all_set$Title <- as.factor(all_set$Title)

# Copying Factors to test dataset
levels(test$Title) <- levels(train$Title)
```

#Setting Age variable for missing values

```{r}
library(rpart)

age_rpart <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title, data = all_set[!is.na(all_set$Age),], method = "anova")

all_set$Age[is.na(all_set$Age)] <- predict(age_rpart, all_set[is.na(all_set$Age),])
```

##Splitting data sets

```{r}
train$Age <- all_set[1:891, "Age"]
train$Title <- all_set[1:891, "Title"]
train$Alias <- all_set[1:891, "Alias"]
test <- all_set[892:1309, ]
```

##Creating a new prediction

```{r}
TitanicRF <- randomForest(as.factor(Survived) ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Age + Title, 
                              data = train, nodesize = 25, ntree = 2444, importance = TRUE)
PredictRF <- predict(TitanicRF, newdata = test)
PredTest <- predict(TitanicRF, newdata=test, type="response")
MySubmission <- data.frame(PassengerID = test$PassengerId, Survived = PredTest)
write.csv(MySubmission, "Submission4.csv", row.names=FALSE)
```

We receive 0.80383 with this submission.

```{r}
varImpPlot(TitanicRF)
```


