---
title: "Kaggle - Analysis with Titatic Data"
author: "Isaac Yauri"
date: "Monday, July 20, 2015"
output: html_document
---

# Titanic: Machine Learning from Disaster

Source: https://www.kaggle.com/c/titanic

Predict survival on the Titanic (using Excel, Python, R, and Random Forests)

See best practice code and explore visualizations of the Titanic dataset on Kaggle Scripts. Submit directly to the competition, no data download or local environment needed!

The sinking of the RMS Titanic is one of the most infamous shipwrecks in history.  On April 15, 1912, during her maiden voyage, the Titanic sank after colliding with an iceberg, killing 1502 out of 2224 passengers and crew. This sensational tragedy shocked the international community and led to better safety regulations for ships.

One of the reasons that the shipwreck led to such loss of life was that there were not enough lifeboats for the passengers and crew. Although there was some element of luck involved in surviving the sinking, some groups of people were more likely to survive than others, such as women, children, and the upper-class.

In this challenge, we ask you to complete the analysis of what sorts of people were likely to survive. In particular, we ask you to apply the tools of machine learning to predict which passengers survived the tragedy.

##VARIABLE DESCRIPTIONS:

survival        Survival
                (0 = No; 1 = Yes)
pclass          Passenger Class
                (1 = 1st; 2 = 2nd; 3 = 3rd)
name            Name
sex             Sex
age             Age
sibsp           Number of Siblings/Spouses Aboard
parch           Number of Parents/Children Aboard
ticket          Ticket Number
fare            Passenger Fare
cabin           Cabin
embarked        Port of Embarkation
                (C = Cherbourg; Q = Queenstown; S = Southampton)

##SPECIAL NOTES:

Pclass is a proxy for socio-economic status (SES)
 1st ~ Upper; 2nd ~ Middle; 3rd ~ Lower

Age is in Years; Fractional if Age less than One (1)
 If the Age is Estimated, it is in the form xx.5

With respect to the family relation variables (i.e. sibsp and parch)
some relations were ignored.  The following are the definitions used
for sibsp and parch.

Sibling:  Brother, Sister, Stepbrother, or Stepsister of Passenger Aboard Titanic
Spouse:   Husband or Wife of Passenger Aboard Titanic (Mistresses and Fiances Ignored)
Parent:   Mother or Father of Passenger Aboard Titanic
Child:    Son, Daughter, Stepson, or Stepdaughter of Passenger Aboard Titanic

Other family relatives excluded from this study include cousins,
nephews/nieces, aunts/uncles, and in-laws.  Some children travelled
only with a nanny, therefore parch=0 for them.  As well, some
travelled with very close friends or neighbors in a village, however,
the definitions do not support such relations.

##We start loading the data sets from Kaggle.

```{r}
setwd("c:/EMC/Cursos/GitHub/Practice-With-Titanic/")
train <- read.csv("train.csv")
test <- read.csv("test.csv")
```

##Loading Libraries

```{r}
library(caTools)
library(ROCR)
library(caret)
library(e1071)
library(randomForest)
library(flexclust)
```

##Review Data

```{r}
summary(train)
summary(test)
```

We have missing values on Age (train and test Dataset) and Fare (test Dataset) variables.

##Working to set missing Fare value for test dataset

We need to set a guess price (Fare) for this observation. Options are: minimun price, maximum price, average price.
But, we can investigate about this person to figure out better what was the Fare.

```{r}
test[which(is.na(test$Fare)), ]
```

Looking prices for people with the same info.

```{r}
MissingFare <- mean(train[train$Pclass == 3 & train$Embarked == "S" & train$Age > 50, "Fare"], na.rm = TRUE)
```

Setting missing Fare with the average of people with the same characteristics.

```{r}
test[153, "Fare"] <- MissingFare
```

##Convertir to Factor for Rain Forrest algorith

```{r}
train$Survived <- as.factor(train$Survived)
train$Pclass <- as.factor(train$Pclass)
train$Embarked <- as.factor(train$Embarked)
test$Pclass <- as.factor(test$Pclass)
test$Embarked <- as.factor(test$Embarked)
```

##Create a Submission.csv file for Kaggle
We receive 0.77990 with this submission.

```{r}
PredTest <- predict(TitanicRF, newdata=test, type="response")
MySubmission <- data.frame(PassengerID = test$PassengerId, Survived = PredTest)
write.csv(MySubmission, "Submission.csv", row.names=FALSE)
```

```{r}
table(train$Survived, train$Sex)
female <- subset(train, Sex == "female")
table(female$Survived, female$Parch)
str(train)
table(train$Parch, train$Survived)
summary(TitanicRF)
table(train$Age)
summary(train$Age)
```

##Trying to solve missing Age info.

To guess what might have been the Age values, we need to:

Set the Age calculating the average of people who have the same caracteristics like: Sex, Class, Fare and Embarked.
We can add a differentiator like Title (Miss, Mrs., Ms., Mme., for example.). To do this, we need to create a new varible and find this characteristic inside the name:

```{r}
head(train$Name)
```

##Creating variable Title

```{r}
train$Title <- NA
test$Title <- NA

train$Title[grep("Master.", train$Name, fixed = TRUE)] <- "Master."
train$Title[grep("Miss", train$Name, fixed = TRUE)] <- "Miss"
train$Title[grep("Mr.", train$Name, fixed = TRUE)] <- "Mr."
train$Title[grep("Mrs.", train$Name, fixed = TRUE)] <- "Mrs."
train$Title[grep("Rev.", train$Name, fixed = TRUE)] <- "Rev."
train$Title[grep("Don.", train$Name, fixed = TRUE)] <- "Don."
train$Title[grep("Dr.", train$Name, fixed = TRUE)] <- "Dr."
train$Title[grep("Major.", train$Name, fixed = TRUE)] <- "Major."
train$Title[grep("Jonkheer", train$Name, fixed = TRUE)] <- "Jonkheer"
train$Title[grep("Col.", train$Name, fixed = TRUE)] <- "Col."
train$Title[grep("Mme.", train$Name, fixed = TRUE)] <- "Mme."
train$Title[grep("Ms.", train$Name, fixed = TRUE)] <- "Ms."
train$Title[grep("Lady.", train$Name, fixed = TRUE)] <- "Lady."
train$Title[grep("Sir.", train$Name, fixed = TRUE)] <- "Sir."
train$Title[grep("Mlle.", train$Name, fixed = TRUE)] <- "Mlle."
train$Title[grep("Capt.", train$Name, fixed = TRUE)] <- "Capt."
train$Title[grep("the Countess.", train$Name, fixed = TRUE)] <- "the Countess"
train$Title[grep("Dona.", train$Name, fixed = TRUE)] <- "Miss"

test$Title[grep("Master.", test$Name, fixed = TRUE)] <- "Master."
test$Title[grep("Miss", test$Name, fixed = TRUE)] <- "Miss"
test$Title[grep("Mr.", test$Name, fixed = TRUE)] <- "Mr."
test$Title[grep("Mrs.", test$Name, fixed = TRUE)] <- "Mrs."
test$Title[grep("Rev.", test$Name, fixed = TRUE)] <- "Rev."
test$Title[grep("Don.", test$Name, fixed = TRUE)] <- "Don."
test$Title[grep("Dr.", test$Name, fixed = TRUE)] <- "Dr."
test$Title[grep("Major.", test$Name, fixed = TRUE)] <- "Major."
test$Title[grep("Jonkheer", test$Name, fixed = TRUE)] <- "Jonkheer"
test$Title[grep("Col.", test$Name, fixed = TRUE)] <- "Col."
test$Title[grep("Mme.", test$Name, fixed = TRUE)] <- "Mme."
test$Title[grep("Ms.", test$Name, fixed = TRUE)] <- "Ms."
test$Title[grep("Lady.", test$Name, fixed = TRUE)] <- "Lady."
test$Title[grep("Sir.", test$Name, fixed = TRUE)] <- "Sir."
test$Title[grep("Mlle.", test$Name, fixed = TRUE)] <- "Mlle."
test$Title[grep("Capt.", test$Name, fixed = TRUE)] <- "Capt."
test$Title[grep("the Countess.", test$Name, fixed = TRUE)] <- "the Countess"
# Dona is in spanich. We need change to Miss.
test$Title[grep("Dona.", test$Name, fixed = TRUE)] <- "Miss"

# Set a factor for this new variable
train$Title <- as.factor(train$Title)
test$Title <- as.factor(test$Title)

# Copying Factors to test dataset
levels(test$Title) <- levels(train$Title)

train[train$Ticket == "PC 17758",]


summary(train)
summary(test)
str(train)
str(test)
table(train$Title, train$Survived)
table(test$Title)
train$Title
table(train$Age, train$Title)
summary(test$Age)      
table(train$Title == "Major.")
train[train$Title == "Master.",]
train[is.na(train$Age)  & train$Title == "Master.",] 
```

```{r}
listacc <- NULL
for (i in (1)) {
    set.seed(2444)
    split <- sample.split(train$Survived, SplitRatio = 0.7)
    Titanictrain <- subset(train, split == TRUE)
    Titanictest <- subset(train, split == FALSE)
    TitanicRF <- randomForest(Survived ~ Pclass + Sex + Fare + Title, 
                              data = Titanictrain, nodesize = 100, ntree = 2444)
    PredictRF <- predict(TitanicRF, newdata = Titanictest)
    tabla <- table(Titanictest$Survived, PredictRF)
    acc <- (tabla[1,1] + tabla[2,2])/nrow(Titanictest)
    #cat("ntree=",i," acc=",acc,"\n")
    listacc[i] <- acc
    if (i %% 100 == 0) { cat(i)}
}
acc
tabla
plot(listacc)
max(listacc)
which.max(listacc)
```

```{r}
PredTest <- predict(TitanicRF, newdata=test, type="response")
MySubmission <- data.frame(PassengerID = test$PassengerId, Survived = PredTest)
write.csv(MySubmission, "Submission.csv", row.names=FALSE)
```


Using K-means clustering

```{r}
k = 2
set.seed(2444)
split <- sample.split(train$Survived, SplitRatio = 0.7)
Titanictrain <- subset(train, split == TRUE)
Titanictest <- subset(train, split == FALSE)
Titanictrain <- Titanictrain[, -6]
summary(Titanictrain)

KMC <- kmeans(Titanictrain$Title, centers = k)
str(KMC)


PredictCART <- predict(titanicCART, newdata = Titanictest, type = "class")
tabla <- table(Titanictest$Survived, PredictCART)
acc <- (tabla[1,1] + tabla[2,2])/nrow(Titanictest)
acc
tabla
prp(TitanicCART)
```

```{r}
PredTest <- predict(TitanicCART, newdata=test, type="response")
MySubmission <- data.frame(PassengerID = test$PassengerId, Survived = PredTest)
write.csv(MySubmission, "Submission.csv", row.names=FALSE)
```

